Squint: Lossy Hierarchical Compression on Symbol Streams

### 2012-12-09 tex/paper.tex
added script for running gen-n tests, more work on results section

\begin{algorithm}
\caption{Similarity Lossifier Algorithm}
\label{sim_alg}
\begin{algorithmic}[1]
\Procedure{Lossify-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \For{$var_1 \rightarrow rhs_1,\ var_2 \rightarrow rhs_2 \in g$}
      \State $str_{i} \gets expand(rhs_i)$
      \If{$lsim(str_{1},str_{2})<\epsilon$}
        \State replace the longer var with the shorter throughout $g$
        \State break
      \EndIf
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

The previous algorithm updates the grammar sequentially, since each
variable replacement can have an effect on large parts of the grammar,
we will have to go through comparing all pairs again after the change.
An alternative algorithm would be to try to replace similar variables in
parallel. This would avoid many of the repeat pair-comparisons done in
\emph{Similarity}. In other words, the algorithm below makes puts 
similar variables
into equivalence class clusters, and then replaces each variable
in a cluster them with a representative
all at once each iteration.

\begin{algorithm}
\caption{Cluster Lossifier Algorithm}
\label{cluster_alg}
\begin{algorithmic}[1]
\Procedure{Cluster-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \State $clusters \gets ()$
    \For{$var_1 \rightarrow rhs_1 \in g$}
      \State blah
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

\section{Results}
\subsection{Illustrative Examples}
It is useful to look at how the algorithm functions on a small piece of
data with both plenty of hierarchical structure, and also some messiness
that most lossless compression schemes would not deal well with. Take
the string below, similar to what one might come across in a math
puzzle book.

@@

Disregarding its meaning, it consists of four very similar equations of the
form $a + b = ?$, and on a smaller scale each term consists of
3 repetitions of approximately the same string of either $123$ or $321$.

Sequitur yields a relatively complicated inferred grammar with 13 rules and
a size of 51. Both the Similarity and Clustering Algorithms with $\epsilon = .4$
, yield the following grammar after reduction. It has 5 rules and is almost
half the size, at 30.

@@

\subsection{Similarity vs Clustering}
For most of our benchmark tests, we focused on two extreme kinds of data:
one with large amounts of repetition at all levels, and one with a word
and sentence level structure but with little obvious large-scale repetition.
One set of test data was constructed by repeating the same letter `a' n-times
but then adding some
texture to the data by changing 1/30 of the characters uniformly
to `*'. Call these test-cases \emph{rep-n}
Another set of test data was constructed from the first n characters of the
book of Genesis from the KJV. Call these test-cases \emph{gen-n}

@@

\begin{table}[t]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
rep-100   & 0.006 & 18/34    & 0.713 & 0.002 & 20/34    & 0.535\\
rep-400   & 0.028 & 45/76    & 0.269 & 0.009 & 41/76    & 0.743\\
rep-1600  & 0.355 & 103/194  & 0.873 & 0.042 & 103/194  & 0.804\\
rep-6400  & 3.978 & 258/448  & 0.706 & 0.203 & 256/448  & 0.869\\
rep-12800
rep-25600 & 135.0 & 900/1482 & 0.829 & 2.906 & 898/1482 & 0.725\\
\end{tabular}
\caption{$\epsilon = .4$}
\end{table}

\begin{table}[t]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
rep-100   & 0.002 & 30/34    & 0.040 & 0.002 & 30/34    & 0.040\\
rep-400   & 0.019 & 60/76    & 0.035 & 0.009 & 60/76    & 0.035\\
rep-1600  & 0.350 & 123/194  & 0.355 & 0.055 & 123/194  & 0.206\\
rep-6400  & 3.942 & 280/448  & 0.387 & 0.295 & 274/448  & 0.543\\
rep-12800  
rep-25600 & 124.5 & 922/1482 & 0.565 & 2.597 & 926/1482 & 0.437\\
\end{tabular}
\caption{$\epsilon = .2$}
\end{table}

TODO: Fill in results for $\epsilon = .1$

@@

TODO: Analogous for genesis
> Growth rate for similarity is far worse for genesis

@@

\subsection{Grammar Reduction Impact}
benchmark two extreme test cases again, with and without reducer
see how much reducer helps (it shouldn't help too much)
notice how slow it is

@@

\section{Conclusions}
\subsection{Ratios and Fidelity}
\subsection{Future Work}
> Better ways to set epsilon to maximize value / cost for different size and types
> Better ways to quantify preservation of hierarchies
> How to find near `self-similarities' that aren't captured by a rule
> Iterating lossify and reduce in a cycle

>>>>

\begin{algorithm}[h]
\caption{Similarity Lossifier Algorithm}
\label{sim_alg}
\begin{algorithmic}[1]
\Procedure{Lossify-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \For{$var_1 \rightarrow rhs_1,\ var_2 \rightarrow rhs_2 \in g$}
      \State $str_{i} \gets expand(rhs_i)$
      \If{$lsim(str_{1},str_{2})<\epsilon$}
        \State replace the longer var with the smaller throughout $g$
        \State break
      \EndIf
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

The previous algorithm updates the grammar sequentially, since each
variable replacement can have an effect on large parts of the grammar,
we will have to go through comparing all pairs again after the change.
An alternative algorithm would be to try to replace similar variables in
parallel. This would avoid many of the repeat pair-comparisons done in
\emph{Similarity}. In other words, the algorithm below puts 
similar variables
into equivalence class clusters, and then replaces each variable
in a cluster them with a representative
all at once each iteration.

\begin{algorithm}[h]
\caption{Cluster Lossifier Algorithm}
\label{cluster_alg}
\begin{algorithmic}[1]
\Procedure{Cluster-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \State $clusters \gets ()$
    \For{$var_1 \rightarrow rhs_1 \in g$}
      \For{$c \in clusters$}
        \State $(var_2 \rightarrow rhs_2) \gets $ first variable in $c$
        \State $str_{i} \gets expand(rhs_i)$
        \If{$lsim(str_{1},str_{2})<\epsilon$}
          \State add $var_1$ to $c$
          \State break
        \EndIf
      \EndFor
      \If{$var_1$ hasn't been added to a cluster}
        \State Add $var_1$ to a new cluster in $clusters$
      \EndIf
    \EndFor
    \For{Cluster $c \in clusters$}
      \State replace all $var \in c$ with the smallest.
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

Both of the algorithms depend on the concept of the "smallest / smaller"
variable. We can define the size of a variable to be the length of
the expansion of its rhs. In this way, if we think of our CFG's as a DAG, then
variables higher up on the DAG will always have larger expansions than
those lower down, and thus will have a larger size. By always replacing
with the smallest variable, we avoid creating cyclic references in our CFG.
We will prove this fact in the next section.

@@

\section{Results}
\subsection{Illustrative Examples}
It is useful to look at how the algorithms function on a small piece of
data with both plenty of hierarchical structure, and also some messiness
that most lossless compression schemes would not deal well with. Take
the string below, similar to what one might come across in a math
puzzle book.

@@

Disregarding its meaning, it consists of four very similar equations of the
form $a + b = ?$, and on a smaller scale each summand consists of
3 repetitions of approximately the same string of either $123,$ or $321,$.

When comparing grammars, we chose to define the \emph{size} of a grammar
as the sum of the lengths of the rhs's of the production rules.
Sequitur yields a relatively complicated inferred grammar with 13 rules and
a size of 51. Both the Similarity and Clustering Algorithms with $\epsilon = .4$
, yield the following grammar after reduction. It has 5 rules and is almost
half the size, at 30.

@@

\subsection{Similarity vs Clustering}
For most of our benchmark tests, we focused on two extreme kinds of data:
one with large amounts of repetition at all levels, and one with a word
and sentence level structure but with little obvious large-scale repetition.
One set of test data was constructed by repeating the same letter `a' n-times
but then adding some
texture to the data by changing 1/30 of the characters uniformly
at random to `*'. Call these test-cases \emph{rep-n}
Another set of test data was constructed from the first n characters of the
book of Genesis from the KJV. Call these test-cases \emph{gen-n}

@@

Below are the results on running Sequitur and Cluster, without
the grammar reductions, on the rep-n datasets. Time is the wall clock
running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
Ratio is the relative size of the grammar produced compared with
Sequitur, where the size of the grammar is the sum of the lengths of the
rhs's of the rules. Levdistance is the levenshtein distance between
the output of the lossified grammar and the original input string, it
is a kind of measure of the fidelity of our algorithms but doesn't
take into account how well patterns were preserved.

\begin{table}[h]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
rep-100   & 0.006 & 18/34    & 0.713 & 0.002 & 20/34    & 0.535\\
rep-400   & 0.028 & 45/76    & 0.269 & 0.009 & 41/76    & 0.743\\
rep-1600  & 0.355 & 103/194  & 0.873 & 0.042 & 103/194  & 0.804\\
rep-6400  & 3.978 & 258/448  & 0.706 & 0.203 & 256/448  & 0.869\\
rep-12800 & & & & & & \\
rep-25600 & 135.0 & 900/1482 & 0.829 & 2.906 & 898/1482 & 0.725\\
\end{tabular}
\caption{rep-n: $\epsilon = .4$}
\end{table}

\begin{table}[h]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
rep-100   & 0.002 & 30/34    & 0.040 & 0.002 & 30/34    & 0.040\\
rep-400   & 0.019 & 60/76    & 0.035 & 0.009 & 60/76    & 0.035\\
rep-1600  & 0.350 & 123/194  & 0.355 & 0.055 & 123/194  & 0.206\\
rep-6400  & 3.942 & 280/448  & 0.387 & 0.295 & 274/448  & 0.543\\
rep-12800 & & & & & & \\
rep-25600 & 124.5 & 922/1482 & 0.565 & 2.597 & 926/1482 & 0.437\\
\end{tabular}
\caption{rep-n: $\epsilon = .2$}
\end{table}

\begin{table}[h]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
rep-100   & & /34    & & & /34    & \\
rep-400   & & /76    & & & /76    & \\
rep-1600  & & /194  & & & /194  & \\
rep-6400  & & /448  & & & /448  & \\
rep-12800 & & & & & & \\
rep-25600 & & /1482 & & & /1482 & \\
\end{tabular}
\caption{$rep-n: \epsilon = .1$}
\end{table}

@@

Repeating the above tests on the \emph{gen-n} datasets yields the following:

\begin{table}[t]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
gen-100   & & /& & & /& \\
gen-400   & & /& & & /& \\
gen-1600  & & /& & & /& \\
gen-6400  & & /& & & /& \\
gen-12800 & & /& & & /& \\
gen-25600 & & /& & & /& \\
\end{tabular}
\caption{$gen-n: \epsilon = .4$}
\end{table}

\begin{table}[t]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
gen-100   & & /& & & /& \\
gen-400   & & /& & & /& \\
gen-1600  & & /& & & /& \\
gen-6400  & & /& & & /& \\
gen-12800 & & /& & & /& \\
gen-25600 & & /& & & /& \\
\end{tabular}
\caption{$gen-n: \epsilon = .2$}
\end{table}

\begin{table}[t]
\begin{tabular}{l|l|l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & S-L(evdistance) & C-T & C-R & C-L \\
\hline
gen-100   & & /& & & /& \\
gen-400   & & /& & & /& \\
gen-1600  & & /& & & /& \\
gen-6400  & & /& & & /& \\
gen-12800 & & /& & & /& \\
gen-25600 & & /& & & /& \\
\end{tabular}
\caption{$gen-n: \epsilon = .1$}
\end{table}

@@

\subsection{Grammar Reduction Impact}
Although the grammars generated by Sequitur are irreducible,
our lossifiers produce grammars which could potentially be
compressed further (losslessly) using Reduction Rules without
changing the output string. To test the cost and
effectiveness of applying reduction rules, we looked at the
rep-n and gen-n test cases again, this time focusing on the
clustering algorithm for $\epsilon = .2$.

\begin{table}[t]
\begin{tabular}{l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & C-T & C-R \\
\hline
rep-100   & & /& & /\\
rep-400   & & /& & /\\
rep-1600  & & /& & /\\
rep-6400  & & /& & /\\
rep-12800 & & /& & /\\
rep-25600 & & /& & /\\
\end{tabular}
\caption{$rep-n: \epsilon = .2$}
\end{table}

\begin{table}[t]
\begin{tabular}{l|l|l|l|l}
Dataset & S-T(ime) & S-R(atio) & C-T & C-R \\
\hline
gen-100   & & /& & /\\
gen-400   & & /& & /\\
gen-1600  & & /& & /\\
gen-6400  & & /& & /\\
gen-12800 & & /& & /\\
gen-25600 & & /& & /\\
\end{tabular}
\caption{$gen-n: \epsilon = .2$}
\end{table}

@@

\section{Conclusions}
\subsection{Algorithm Tradeoffs}
Runtimes for each algorithm
Considerations in choosing epsilon
\subsection{Ratios and Fidelity}
\subsection{Future Work}
* Better ways to set epsilon to maximize value / cost for different size and types
* Better ways to quantify preservation of hierarchies
* How to find near `self-similarities' that aren't captured by a rule
* Iterating lossify and reduce in a cycle
